What We Eval in the Shadows

### 2021-04-14 paper/main.tex
Example of match.call and model.frame

% We have quantitative data on that if needed...
This pattern is very often (more than half of the cases) used in relation with statistical models, to build a new model. \c{model.frame} to build a new generic model, or a more specialized model like \c{glm} or, \c{lm}.
% Show example

@@

 A call to \c{library(packageName)}, which loads libraries in R, for instance, will
populate environment \c{package:packageName} using \eval. It is a \emph{base} function
though, so we do not see it in the CRAN dataset. However, R still makes it
possible to access any package environments and \packageNbPackageEnvSites call
sites, \ie \packagePackageEnvSitePercent access a \emph{package} environment, from
\packageNbPackageEnvPackages different packages. Those packages perform
introspection, to represent and search any environment by name, in package
\emph{envnames}, to containerize tests, as with \emph{testthat} and
\emph{unitizer}, to create a custom OOP system with \emph{R.oo}, or to bypass the visibility of functions in \emph{USSCXenaTools}. It is also used to
implement a custom caching mechanism in packages \emph{g.data} and \emph{SOAR}.

>>>>

% We have quantitative data on that if needed...
This pattern is very often (more than half of the cases) used in relation with statistical models, to build a new model, \c{model.frame} to build a new generic model, or a more specialized model like \c{glm} or, \c{lm}. The R documentation of \c{match.call} explicitly describes this usage.

In package \emph{survival}\footnote{It is a \emph{recommended package}, so recommended by the CRAN maintainers, thoroughly tested, and following R best practices.}, in function \c{coxph}, the surrounding call is first captured, then \c{model.frame} is injected as the function name in the call, and finally, the new call is evaluated in the parent frame.

\begin{lstlisting}
Call <- match.call()
...
tform <- Call[c(1,indx)]  # only keep the arguments we wanted
tform[[1L]] <- quote(stats::model.frame)  # change the function called
...
mf <- eval(tform, parent.frame())
\end{lstlisting}

@@

 A call to \c{library(packageName)}, which loads libraries in R, for instance, will
populate environment \c{package:packageName} using \eval. It is a \emph{base} function
though, so we do not see it in the CRAN dataset. However, R still makes it
possible to access any package environments, which we see in
\packageNbPackageEnvPackages different packages in our corpus. Those packages perform
introspection, to represent and search any environment by name, in package
\emph{envnames}, to containerize tests, as with \emph{testthat} and
\emph{unitizer}, to create a custom OOP system with \emph{R.oo}, or to bypass the visibility of functions in \emph{USSCXenaTools}. It is also used to
implement a custom caching mechanism in packages \emph{g.data} and \emph{SOAR}.
